---
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---
# The damage of different weather events on public health and economy
###Abstract:
    we analyze the impact of different weather events on public health and economy in the report. We use the four attributes of fatalities, injuries, property and crop damage to estate which types of event are most harmful to the population health and economy, on the basis of analysis,we can conclude that excessive heat and tornado are more harmful to population health, and flood, drought, and hurricane/typhoon have the worst economic consequences.

```{r,echo=TRUE}
 #set the directory and  require  the packages 
setwd("E:/")  
library(data.table)
library(ggplot2)
library(gridExtra)
```

```{r,results='hide'}
#get the origin datasets and manipulate
data<-fread("data.csv")
```

```{r,results='hide'}

data$year <- as.numeric(format(as.Date(data$BGN_DATE, format = "%m/%d/%Y %H:%M:%S"), "%Y"))
ggplot(data,aes(year))+geom_histogram()   #show the counts of years
```

```{r,echo=TRUE}
# get the selected datasets we use
year<-as.data.frame(table(data$year))
year$cum<-cumsum(year$Freq)
year
dim(data)[1]*0.8  #to compare with the cum attributes in the year 
storm<-data[data$year>=1992,]   #after comparing ,we use the data after 1992,the datasets of storm is what we use next,its class is  also data.table
```

```{r}
#the Fatalities by Severe Weather  from 1992 to 2011
fata<-storm[,list(all.fata=sum(FATALITIES)),by=EVTYPE]
fata<-as.data.frame(fata)
fata_top15<-fata[order(fata$all.fata,decreasing =T),][1:15,]
fata_top15
p1<-ggplot(fata_top15, aes(x=EVTYPE, y=all.fata)) + 
    geom_bar(stat ="identity",width=0.8)+ 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    xlab("Severe Weather Type") + ylab("Number of Fatalities") + 
    ggtitle("Total Fatalities by Severe Weather\n Events in the U.S. from 1992 - 2011")
```

```{r}
#the Injuries by Severe Weather  from 1992 to 2011
inj<-storm[,list(all.inj=sum(INJURIES)),by=EVTYPE]
inj<-as.data.frame(inj)
inj_top15<-inj[order(inj$all.inj,decreasing =T),][1:15,]
inj_top15
p2<-ggplot(inj_top15, aes(x=EVTYPE, y=all.inj)) + 
    geom_bar(stat ="identity",width=0.9)+ 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    xlab("Severe Weather Type") + ylab("Number of Injuries") + 
    ggtitle("Total Injuries by Severe Weather\n Events in the U.S. from 1992 - 2011")
```

```{r}
grid.arrange(p1,p2,ncol=2)
```

```{r}
#the Property damage by Severe Weather  from 1992 to 2011
a<-c("EVTYPE","PROPDMG","PROPDMGEXP","CROPDMG","CROPDMGEXP")
storm<-as.data.frame(storm)
data2<-storm[,names(storm) %in% a]
data2[data2$PROPDMGEXP=="",]$PROPDMGEXP=0
data2[tolower(data2$PROPDMGEXP)=="h",]$PROPDMGEXP=2
data2[tolower(data2$PROPDMGEXP)=="k",]$PROPDMGEXP=3
data2[tolower(data2$PROPDMGEXP)=="m",]$PROPDMGEXP=6
data2[tolower(data2$PROPDMGEXP)=="b",]$PROPDMGEXP=9
data2$PROPDMGEXP<-as.numeric(data2$PROPDMGEXP)
data2<-data2[complete.cases(data2),]
data2$PRO<-data2$PROPDMG*10^(data2$PROPDMGEXP)
data2<-as.data.table(data2)

pro<-data2[,list(all.pro=sum(PRO)),by=EVTYPE]
pro<-as.data.frame(pro)
pro_top15<-pro[order(pro$all.pro,decreasing =T),][1:15,]
pro_top15
p3<-ggplot(pro_top15, aes(x=EVTYPE, y=all.pro)) + 
    geom_bar(stat ="identity",width=0.9)+ 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    xlab("Severe Weather Type") + ylab("Number of Property damage") + 
    ggtitle("Total Property damage by Severe Weather\n Events in the U.S. from 1992 - 2011")
```

```{r}
#the crop damage by Severe Weather  from 1992 to 2011
a<-c("EVTYPE","PROPDMG","PROPDMGEXP","CROPDMG","CROPDMGEXP")
storm<-as.data.frame(storm)
data3<-storm[,names(storm) %in% a]
data3[data3$CROPDMGEXP=="",]$CROPDMGEXP=0
data3[tolower(data3$CROPDMGEXP)=="k",]$CROPDMGEXP=3
data3[tolower(data3$CROPDMGEXP)=="m",]$CROPDMGEXP=6
data3[tolower(data3$CROPDMGEXP)=="b",]$CROPDMGEXP=9
data3$CROPDMGEXP<-as.numeric(data3$CROPDMGEXP)
data3<-data3[complete.cases(data3),]
data3$crop<-data3$CROPDMG*10^(data3$CROPDMGEXP)
data3<-as.data.table(data3)

crop<-data3[,list(all.crop=sum(crop)),by=EVTYPE]
crop<-as.data.frame(crop)
crop_top15<-crop[order(crop$all.crop,decreasing =T),][1:15,]
crop_top15

p4<-ggplot(crop_top15, aes(x=EVTYPE, y=all.crop)) + 
    geom_bar(stat ="identity",width=0.9)+ 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    xlab("Severe Weather Type") + ylab("Number of Crop damage") + 
    ggtitle("Total Crop damage by Severe Weather\n Events in the U.S. from 1992 - 2011")
```

```{r}
grid.arrange(p3,p4,ncol=2)
```
